---
title: "Brownlows Bets"
author: "Jeff Roderick"
date: "10-16-19"
output: html_document: default
 ---

<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

 At the first day of class there are 35 students in Math Statistics class and professor Brownlow always proposes this bet to his students "I will wager $2 to your bet of $1 that at least two people in today's class have  the same birthday.  He tells me he has made that bet 1137 times so far. Where do you think his profits are? Discuss this in a markdown document

# Methods

<!--Decide on your methods:  In this section explain the approach you will take to attack the issue --- how will you decide on a likely estimate of professor Brownlow's earnings or losses after 1137 bets. Also explain how you construct a 95% confidence interval about your estimate of his winnings !-->
The way I will do is by creating a data set, run a loop 1137 times in a sample, and it will allow use to see if there two of the same days in a month that are the same using the unqiue function.
Based on if he wins the bet he gains $1.00 and if he loses a bet, he will lose $2.00. We we will get a confindence interval by creating histogram. On this histogram with each tail will be 5% and the rest will get us the 95%.


# Profit Estimate

<!--Explain and display the results of your method of coming up with the estimate of professor Brownlows betting!-->

First we create the year data and a vector to contain.
```{r}
Bdays = 1:365
Bsamples = vector("numeric")
TotalDollars= 0
saveDollars=vector('numeric')
```

run the simulation 1137 times

```{r}
for(loop in 1:1000){

  TotalDollars= 0
  classSize= 35
  for(i in 1:1137) {
    Bsamples = sample(Bdays,35, replace= TRUE)
   if (length(unique(Bsamples)) == classSize) {
     TotalDollars = TotalDollars -2
   } else{
      TotalDollars = TotalDollars +1
  }  
 }

saveDollars[loop] = TotalDollars
}
hist(saveDollars)
print(mean(saveDollars))





```




# Confidence interval of Brownlow's earning

<!--Explain and display your method of coming up with a 95% confidence interval of  professor Brownlows-->
We assign values for mean, standard deviation, and sample size to a rstudio in a formula.
We use our data to create a code to provide us with a confidence interval. 
```{r}
stdevDollars = sd(saveDollars)
meanDollars = mean(saveDollars)
CI = c(meanDollars-2*stdevDollars, meanDollars+2*stdevDollars)

print(CI)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If professor Brownlow is consistantly making or losing money how could he change his payout so that his profits center on zero?  Also discuss what size his class should be in order to have his current scheme balance at zero profit.!-->

```{r}
TotalDollars= 0
saveDollars=vector('numeric')
for(loop in 1:1000){

  TotalDollars= 0
  classSize= 29
  for(i in 1:1137) {
    Bsamples = sample(Bdays,classSize, replace= TRUE)
   if (length(unique(Bsamples)) == classSize) {
     TotalDollars = TotalDollars -2
   } else{
      TotalDollars = TotalDollars +1
  }  
 }

saveDollars[loop] = TotalDollars
}
hist(saveDollars)
print(mean(saveDollars))





```


```{r}
TotalDollars= 0
saveDollars=vector('numeric')
classSize= 35
lossAmt = 4.5

for(loop in 1:100){

  TotalDollars= 0
 
  for(i in 1:1137) {
    Bsamples = sample(Bdays,classSize, replace= TRUE)
   if (length(unique(Bsamples)) == classSize) {
     TotalDollars = TotalDollars - lossAmt
   } else{
      TotalDollars = TotalDollars +1
  }  
 }

saveDollars[loop] = TotalDollars
}
hist(saveDollars)
print(mean(saveDollars))





```

 If his class size 29 then his payout is around zero profit, which is his goal. The payout to break even must be $4.50 to one. It appear this gentlemn may have a bad buisness establishment, if he plans on making a zero profit. I ran the data above with addition r chunks to display coding and answers above to my reasoning. 
